---
title: "HW2_script"
author: "Christina Treacy"
date: "May 7, 2026"
format: pdf
editor: visual
---

```{r}
#install.packages("dataRetrieval")
library(dataRetrieval)
library(tidyverse)
library(here)
```

Load in data from NWIS server and save to CSV

```{r}
siteNumbers <- c("14211720", "14211010", "14166000")
parameterCd <- "00010" # water temperature
startDate <- "2010-01-01"
endDate <- "2023-12-31"
temp_data <- readNWISdv(siteNumbers, parameterCd, startDate, endDate)

# rename columns for clarity
colnames(temp_data) <- c("agency_cd", "site_no", "Date", "temp_C")
# extract some date info for use later
temp_data$month = month(temp_data$Date, label=TRUE)
temp_data$year = year(temp_data$Date)

write.csv(temp_data, "tempdata.csv", row.names = FALSE)
```

Load data directly from repository

```{r}
here()

tempdata <- read.csv(here("tempdata.csv"))
```

Plot data & Calculate Summary Statistics
```{r}
# check by plotting
ggplot(tempdata, aes(x=Date, y=temp_C, color=site_no)) +
geom_line() +
labs(title="Stream Temperature at Willamette River Sites",
x="Date", y="Temperature (°C)", color="Site Number") +
theme_minimal()

# generate summary stats
summary_table <- tempdata |>
  group_by(site_no) |>
  summarize(min=min(temp_C), 
            mean=mean(temp_C),
            med=median(temp_C),
            max=max(temp_C),
            st_dev=sd(temp_C)) 
write.csv(summary_table, here("site_temp_summary_table.csv"))
```

Trends in Water Temperature
a) is there a trend? 
b) is the slope (trend) positive (increasing) or negative (decreasing)?
c) are there differences between sites?

```{r}
# filter for August temperatures
aug_data <- subset(temp_data, month == "Aug")

# fit a GLM model to see if there is a trend in August temperatures over time, separated by sit
aug_model <- glm(temp_C ~ year*site_no, data = aug_data)
summary(aug_model)
```

Estimate for year tells you slope for first site: 0.12975

Estimate for year:site_no14211010 (0.04627), year:site_no14211720 (0.01619) tells you how much greater (if positive) slope is for this site relative to first site: 

Estimates for Intercept (-243.03103) and site_no14211010 (-91.98714), site_no14211010 (-28.41636) set the baseline temperature to make regression work. You won’t use these. (An easier way to look at differences in mean temperature across sites is to calculate means for each)


