---
title: "HW2_script"
author: "Christina Treacy"
date: "May 7, 2026"
format: pdf
editor: visual
---

```{r}
#install.packages("dataRetrieval")
library(dataRetrieval)
library(tidyverse)
library(here)
```

Load in data from NWIS server and save to CSV

```{r}
siteNumbers <- c("14211720", "14211010", "14166000")
parameterCd <- "00010" # water temperature
startDate <- "2010-01-01"
endDate <- "2023-12-31"
temp_data <- readNWISdv(siteNumbers, parameterCd, startDate, endDate)

# rename columns for clarity
colnames(temp_data) <- c("agency_cd", "site_no", "Date", "temp_C")
# extract some date info for use later
temp_data$month = month(temp_data$Date, label=TRUE)
temp_data$year = year(temp_data$Date)

write.csv(temp_data, "tempdata.csv", row.names = FALSE)
```

Load data directly from repository

```{r}
here()

tempdata <- read.csv(here("tempdata.csv"))
```

Plot data & Calculate Summary Statistics
```{r}
# check by plotting
ggplot(temp_data, aes(x=Date, y=temp_C, color=site_no)) +
geom_line() +
labs(title="Stream Temperature at Willamette River Sites",
x="Date", y="Temperature (°C)", color="Site Number") +
theme_minimal()

# generate summary stats

```

```{r}
# filter for August temperatures
feb_data <- subset(temp_data, month == "Feb")

# fit a GLM model to see if there is a trend in August temperatures over time, separated by sit
feb_model <- glm(temp_C ~ year*site_no, data = feb_data)
summary(feb_model)

```

Estimate for year tells you slope for first site:  -0.064277

Estimate for year:site_no14211010 (0.019204), year:site_no14211720 (0.004215) tells you how much greater (if positive) slope is for this site relative to first site 

Estimates for Intercept (136.647423) and site_no14211010 (-40.061916), site_no14211010 (-8.440431) set the baseline temperature to make regression work. You won’t use these. (An easier way to look at differences in mean temperature across sites is to calculate means for each)